\trtitle

### 2025-01-21 BAI/paper/presentation.tex
improved presentation a little bit.

% Slide 1: Introduction
\begin{frame}{Introduction}
	\textbf{What is Image Captioning?}
	\begin{itemize}
			\item Task: Generate descriptive textual captions for images.
			\item Combines computer vision (CV) and natural language processing (NLP).
			\item Applications: Assistive technology, automated metadata generation, and more.
	\end{itemize}
	% Speaker Notes:
	% Begin by briefly explaining the task of image captioning. Highlight its interdisciplinary nature and practical applications.
\end{frame}

% Slide 2: Problem Statement
\begin{frame}{Problem Statement}
	\textbf{Challenges in Image Captioning}
	\begin{itemize}
			\item Detecting and interpreting visual elements accurately.
			\item Generating coherent and contextually relevant text.
			\item Balancing computational efficiency and model performance.
	\end{itemize}
	% Speaker Notes:
	% Emphasize the dual challenges of visual understanding and text generation. Mention the need for scalable solutions in real-world applications.
\end{frame}

% Slide 3: Methodology (Dataset and Features)
\begin{frame}{Methodology: Dataset and Features}
	\textbf{Dataset: Flickr8k}
	\begin{itemize}
			\item Contains 8,000 images with five captions per image.
	\end{itemize}
	\textbf{Feature Extraction}
	\begin{itemize}
			\item Pre-trained EfficientNet B0 model.
			\item Extracts high-level features from images.
	\end{itemize}
	% Speaker Notes:
	% Explain why Flickr8k was chosen and describe its characteristics. Introduce EfficientNet B0 as a state-of-the-art feature extractor.
\end{frame}

% Slide 4: Methodology (Decoder Architectures)
\begin{frame}{Methodology: Decoder Architectures}
	\textbf{Architectures Evaluated}
	\begin{itemize}
			\item GRU (Gated Recurrent Units)
			\item LSTM (Long Short-Term Memory Networks)
			\item Transformer
	\end{itemize}
	\textbf{Implementation}
	\begin{itemize}
			\item Implemented in PyTorch.
			\item Trained using image features and paired captions.
	\end{itemize}
	% Speaker Notes:
	% Briefly describe each architecture and its role in the pipeline. Mention the choice of PyTorch for implementation.
\end{frame}

% Slide 5: Results (Metrics)
\begin{frame}{Results: Metrics}
	\textbf{Evaluation Metrics}
	\begin{itemize}
			\item BLEU: Measures n-gram overlap.
			\item METEOR: Considers semantic matching.
			\item NIST: Focuses on information content.
	\end{itemize}
	\textbf{Key Findings}
	\begin{itemize}
			\item Transformer underperforms GRU and LSTM in all metrics.
	\end{itemize}
	% Speaker Notes:
	% Introduce the metrics and explain their relevance. Highlight the surprising result of Transformer underperformance.
\end{frame}

% Slide 6: Results (Comparison)
\begin{frame}{Results: Comparison of Architectures}
	\begin{itemize}
			\item GRU and LSTM outperform Transformer.
			\item Transformer struggles with smaller datasets like Flickr8k.
	\end{itemize}
	\textbf{Visual Comparison}
	\includegraphics[width=.6\textwidth]{res/metrics.png}
	% Speaker Notes:
	% Present the comparison visually and discuss why the Transformer might struggle with limited data.
\end{frame}

% Slide 7: Discussion
\begin{frame}{Discussion}
	\textbf{Insights and Implications}
	\begin{itemize}
			\item Traditional architectures (GRU, LSTM) remain competitive.
			\item Transformers may require larger datasets to realize their full potential.
	\end{itemize}
	% Speaker Notes:
	% Reflect on what these results mean for the field and the practical implications of using different architectures.
\end{frame}

% Slide 8: Conclusion and Future Work
\begin{frame}{Conclusion and Future Work}
	\textbf{Summary}
	\begin{itemize}
			\item Evaluated GRU, LSTM, and Transformer for image captioning.
			\item GRU and LSTM outperform Transformer on Flickr8k.
	\end{itemize}
	\textbf{Future Work}
	\begin{itemize}
			\item Test on larger datasets (e.g., MS COCO).
			\item Explore hybrid architectures.
	\end{itemize}
	% Speaker Notes:
	% Summarize the study's key points and suggest future directions for improvement and exploration.
\end{frame}

>>>>

\begin{frame}{Introduction}
	\textbf{What is Image Captioning?}
	\begin{itemize}
			\item Task: Generate descriptive textual captions for images.
			\item Combines computer vision (CV) and natural language processing (NLP).
			\item Application examples: Assistive technology, automated metadata generation.
	\end{itemize}
\end{frame}

\begin{frame}{Introduction}
	\begin{figure}[H]
		\includegraphics[width=.6\textwidth]{res/dog.jpg}
		\caption{A dog running on a field.}\label{fig:dog}
	\end{figure}
\end{frame}

\begin{frame}{Research Question}
	How do different decoder architectures (GRU, LSTM, Transformer) perform in the context of image captioning on the Flickr8k dataset?
\end{frame}

\begin{frame}{Methodology: Dataset and Features}
	\textbf{Dataset: Flickr8k}
	\begin{itemize}
			\item Contains 8,000 images with five captions per image.
	\end{itemize}
	\textbf{Feature Extraction}
	\begin{itemize}
			\item Pre-trained EfficientNet B0 model.
			\item Extracts high-level features from images.
	\end{itemize}
\end{frame}

\begin{frame}{Methodology: Decoder Architectures}
	\textbf{Architectures Evaluated}
	\begin{itemize}
			\item GRU (Gated Recurrent Units)
			\item LSTM (Long Short-Term Memory Networks)
			\item Transformer
	\end{itemize}
	\textbf{Implementation}
	\begin{itemize}
			\item Implemented in PyTorch.
			\item Trained using image features and paired captions.
	\end{itemize}
\end{frame}

\begin{frame}{Methodology: Decoder Architectures}
	\begin{figure}[H]
		\includegraphics[width=.9\textwidth]{res/gru.png}
		\caption{GRU Decoder.}\label{fig:gru}
	\end{figure}
\end{frame}

\begin{frame}{Methodology: Decoder Architectures}
	\begin{figure}[H]
		\includegraphics[width=.9\textwidth]{res/lstm.png}
		\caption{LSTM Decoder.}\label{fig:lstm}
	\end{figure}
\end{frame}

\begin{frame}{Methodology: Decoder Architectures}
	\begin{figure}[H]
		\includegraphics[width=.9\textwidth]{res/transformer.png}
		\caption{Transformer Decoder.}\label{fig:transformer}
	\end{figure}
\end{frame}

\begin{frame}{Results: Metrics}
	\textbf{Evaluation Metrics}
	\begin{itemize}
			\item BLEU: Measures n-gram overlap.
			\item METEOR: Considers semantic matching.
			\item NIST: Focuses on information content.
	\end{itemize}

	% TODO: explain these metrics
\end{frame}

\begin{frame}{Results: Comparison of Architectures}
	\begin{itemize}
			\item GRU and LSTM outperform Transformer.
			\item Transformer struggles with smaller datasets like Flickr8k.
	\end{itemize}
	\textbf{Visual Comparison}
	\includegraphics[width=.6\textwidth]{res/metrics.png}
\end{frame}

\begin{frame}{Discussion}
	\textbf{Insights and Implications}
	\begin{itemize}
			\item Traditional architectures (GRU, LSTM) remain competitive.
			\item Transformers may require larger datasets to realize their full potential.
	\end{itemize}
\end{frame}

\begin{frame}{Conclusion and Future Work}
	\textbf{Summary}
	\begin{itemize}
			\item Evaluated GRU, LSTM, and Transformer for image captioning.
			\item GRU and LSTM outperform Transformer on Flickr8k.
	\end{itemize}
	\textbf{Future Work}
	\begin{itemize}
			\item Test on larger datasets (e.g., MS COCO).
			\item Finetune pretrained models.
	\end{itemize}
\end{frame}
